---
title: "Gov 50 Final Project"
author: "Abby Loeb"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

```{r}
library(tidyverse)
library(ggplot2)
averages <- read_csv("averages.csv")
```

## Dataset description

Does being elected from a swing state change how often a congressional member votes in alignment with Trump? I plan to study how often various representatives vote in alignment with Donald Trump's stated policy positions. I hypothesize that congressional members from swing states will vote closer to the median Trump vote, for both parties relative to the overall party median Trump vote. I expect that these members will, on average, vote closer to the overall median in order to win re-election in a competitive seat where a member must apply to undecided or voters who feel as though they are in the middle of each party ideologically. I'll be looking at the `state` and `net_trump_vote` variables of each member. The `net_trump_vote` is measured on a scale of -1 to 1, where -1 are more votes against Trump. I also may group these members by `state` and `party` and then analyze the `net_trump_vote` across these groups to get a better understanding of how a member's party and and state effect their voting patterns. I expect that the difference in the `net_trump_vote` between the two `party` groups within a singular swing state will be less than the respective `party` `net_trump_vote` in non-swing states. My hypothesis would be disproven if there was no considerable difference in the difference of the `net_trump_vote` between parties in swing states versus non-swing states.



## Milestone 3

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
averages <- read_csv("averages.csv")
averages
```
```{r}
average_agree <- averages |>
  group_by(state) |>
  summarize(avg_agree_pct = mean(agree_pct))

average_agree |>
  slice_max(avg_agree_pct, n = 50) |>
  ggplot(mapping = aes(x = avg_agree_pct, y = state)) +
  geom_col() +
  labs(
    x = "Average vote percent with Trump by state",
    y = "State",
    title = "Average agreement with Trump by state"
  )
```

These graphs show no obvious pattern based on state alone, with the average Trump vote aligning with the overall political tilt of the specific state. More Democratic states have representatives who tend to vote less with Trump on average, and Republican-leaning states are represented by House members who tend to vote more with Trump on average.

This makes clear that we need to further break down the data in swing states by party and/or member to understand how members from swing states vote differently from other members of their party and whether this correlates with being elected by a district in a swing state or not.

## Milestone 4

```{r}
averages <- averages |>
  mutate(swing = case_when(
    state == "FL" ~ "Swing State", 
    state == "NC" ~ "Swing State", 
    state == "NV" ~ "Swing State", 
    state == "PA" ~ "Swing State",
    state == "MI" ~ "Swing State",
    state == "WI" ~ "Swing State",
    state == "AZ" ~ "Swing State",
    state == "NH" ~ "Swing State",
    state == "IA" ~ "Swing State",
    state == "GA" ~ "Swing State",
    state == "MN" ~ "Swing State",
    state == "CO" ~ "Swing State",
    state == "TX" ~ "Swing State",
    state == "OH" ~ "Swing State",
    state == "VA" ~ "Swing State",
  )) |>
  mutate(swing = if_else(is.na(swing), "Not Swing State", "Swing State")) |>
  mutate(party = case_when(
    party == "D" ~ "Democrat",
    party == "Democrat" ~ "Democrat",
    party == "R" ~ "Republican",
    party == "Republican"~"Republican",
    party == "I" ~ "Independent",
    party == "Independent" ~ "Independent",
    party == "Libertarian" ~ "Libertarian"
  )) |>
  drop_na(net_trump_vote, party)
averages <- averages |>
  subset(party == "Democrat" | party == "Republican")
``` 

In this chunk, I first made a new variable which denotes whether a state is a swing state or not swing state by the standards of the 2020 Presidential election. For this, I used the states that CNN Politics considered battlegrounds, regardless of margin. I then mutated the party names to standardize them, as some were full party names and some were abbreviated. I dropped all rows without a net Trump vote score, as that is the variable of interest. Finally, I subsetted this data to only include Democrats and Republicans, as both Independents and Libertarians had too few representatives, and these amounts fluctuated across congresses and swing states/not swing states, that I couldn't conclude much from them.

I am using the net Trump vote of each legislator to gather my conclusions. This is the difference of the amount of times a legislator in the House has voted with and against Trump's position on legislation.

```{r}
DiM_party<-averages |>
  group_by(party) |>
  summarize(mean_ntv = mean(net_trump_vote)) |>
  pivot_wider(
    names_from = party,
    values_from = mean_ntv
  )

knitr::kable(DiM_party, col.names = c("Avg. net Trump vote, Democrats", "Avg. net Trump vote, Republicans"))

DiM_swing <- averages |>
  group_by(swing, party) |>
  summarize(mean_ntv = mean(net_trump_vote)) |>
  pivot_wider(
    names_from = swing,
    values_from = mean_ntv
  ) |>
  mutate(ntv_diff = `Swing State` - `Not Swing State`)
DiM_swing
knitr::kable(DiM_swing, col.names = c("Party", "Avg. Trump Vote, non-swing states", "Avg. Trump Vote, swing states", "Net Trump Vote Difference"))
```

In the first tibble, which is the 1x2 knitted tibble, I measured the average net Trump vote across all legislators of the Democratic and Republican parties. The average net Trump vote is -26.12 across all Democratic legislators, while the net Trump vote for Republican legislators is 22.17.

In the second tibble, which is 2x4, candidates of each party are divided on whether they are from a swing or not swing state, with the difference between these groups reported as well. In this, we can see that Democrats in swing states, with an average net Trump vote of -20.97, is 8.15 units lower than those not frmo swing states who had an average of -29.12. We see a similar pattern among Republicans, as those in swing states had an average net Trump vote of 18.6, 6.76 units lower than the average net Trump vote of Republicans not from swing states, who had an average Trump vote of 25.37. 

In each party, we see that Democrats and Republicans from swing states have an average net Trump vote that is closer to 0, or a net Trump vote score that signifies equally voting with or against the Trump policy on a piece of legislation. This may be due to a desire to appeal to voters that are more ready to switch or run in districts that are decided by a smaller margin or pivotal group of swing voters.


```{r}
averages |>
  group_by(swing, party) |>
  summarize(mean_ntv = mean(net_trump_vote)) |>
  ggplot(aes(x = party, y = mean_ntv, fill = swing)) +
  geom_col() +
  scale_fill_manual(values = c(`Swing State` = "goldenrod1", `Not Swing State` = "goldenrod")) +
  labs(
    x = "Party of Officeholder",
    y = "Mean number of votes in accordance with Trump",
    title = "Average votes in accordance with Trump for 115-117 Congress",
    caption = "Source: FiveThirtyEight"
  )

```

This graph is a visualization of my previous tibble measuring the net Trump vote for each major party across the swing states and not swing states, demonstrating how much closer to 0 the representatives from swing states tend to vote over members of the same party not from swing states.

I think much of this may be affected by national environment, including which Congress they each data point is from, as they are from the 115th, 116th, and 117th Congresses. My next steps would be to examine the data similarly across each Congress.





